Think deep for reef resilience

### 2025-08-14 paper/paper_draft.qmd
Collapse bioregion visual results into a single paragraph
Add methods detail to reflect processing of data from NOAA methodology

Degree Heating Weeks (DHWs) were then derived using the "HotSpot SST Anomaly" approach
[@skirving2020]. The "HotSpot SST Anomaly" represents the thermal stress experienced by
coral reefs and is calculated as the difference between daily SST values and the maximum of
a pixel's monthly mean climatology across all twelve months (i.e., the maximum monthly mean
SST). This approach uses the annual maximum temperature as the baseline because corals are
adapted to withstand their local warmest conditions under normal seasonal cycles. When
temperatures exceed this seasonal maximum, corals experience thermal stress that can lead to
bleaching. Only positive anomalies are considered relevant for coral stress, as temperatures
below the seasonal maximum do not contribute to bleaching risk.

DHW values on a per-pixel basis at 1km^2^ resolution were then calculated as the 12-week
rolling sum of positive HotSpot SST Anomaly values, expressed in units of °C-weeks. The
12-week accumulation period reflects the cumulative nature of coral thermal stress. Brief
temperature spikes may be tolerable, but sustained or repeated exposure to elevated
temperatures causes progressively greater physiological stress and bleaching severity. DHW
values integrate both the magnitude and duration of thermal stress: a moderate anomaly
(e.g., 1°C) sustained for 12 weeks produces the same DHW value as a severe anomaly (e.g.,
3°C) lasting 4 weeks, though the biological impacts may differ. DHW estimates for each
indicative reef (see @GBRMPA2022) were then determined by taking the mean DHW value for all
pixels touching each reef outline. The process described here was conducted for all 2429
reefs to create DHW projections spanning from 2025 to 2075. Code used to calculate DHW data
and perform analyses involved in this investigation can be found at **\[insert repo
citation\]**.

@@

In many bioregions reefs in the 'low' cover cluster have the lowest cover - DHW timeseries
correlation levels, while reefs in 'high' clusters have the highest correlation levels
(@sfig-access-esm1-5-bioregion-dhwcor).

Conversely, there are no consistent patterns across bioregions in reef area-relative cover
and mean timeseries DHW levels, with clusters having similar mean DHW levels in many
bioregions (@sfig-access-esm1-5-bioregion-dhwmean).

Similarly, there are no substantial differences between reef clusters in mean thermal
tolerance levels (@sfig-access-esm1-5-bioregion-dhwtol).

There are no consistent differences across bioregions and GCMs between clusters in their
reefs' initial proportion of habitable area occupied by corals
(@sfig-access-esm1-5-bioregion-initialproportion).

Across most bioregions there are not consistent differences between clusters in connectivity
metrics such as source-to-sink connectivity ratio (@sfig-access-esm1-5-bioregion-logsosi)
and total connectivity strength (@sfig-access-esm1-5-bioregion-totalconnectivity).

>>>>

Degree Heating Weeks (DHWs) were then derived using the "HotSpot SST Anomaly" approach
[@skirving2020]. The "HotSpot SST Anomaly" represents the thermal stress experienced by
coral reefs and is calculated as the difference between daily SST values and the maximum of
a pixel's monthly mean climatology across all twelve months (i.e., the maximum monthly mean
SST). This approach uses the annual maximum temperature as the baseline because corals are
adapted to withstand their local warmest conditions under normal seasonal cycles. When
temperatures exceed this seasonal maximum, corals experience thermal stress that can lead to
bleaching. Only positive anomalies, greater than 1°C, are considered relevant for coral
stress, as temperatures below the seasonal maximum do not contribute to bleaching risk.

DHW values on a per-pixel basis at 1km^2^ resolution were then calculated as the 12-week
rolling sum of positive HotSpot SST Anomaly values that are greater than 1°C, expressed in
units of °C-weeks. The 12-week accumulation period reflects the cumulative nature of coral
thermal stress. Brief temperature spikes may be tolerable, but sustained or repeated
exposure to elevated temperatures causes progressively greater physiological stress and
bleaching severity. DHW values integrate both the magnitude and duration of thermal stress:
a moderate anomaly (e.g., 1°C) sustained for 12 weeks produces the same DHW value as a
severe anomaly (e.g., 3°C) lasting 4 weeks, though the biological impacts may differ. DHW
estimates for each indicative reef (see @GBRMPA2022) were then determined by taking the mean
DHW value for all pixels touching each reef outline. The process described here was
conducted for all 2429 reefs to create DHW projections spanning from 2025 to 2075. Code used
to calculate DHW data and perform analyses involved in this investigation can be found at
**\[insert repo citation\]**.

@@

In many bioregions reefs in the 'low' cover cluster have the lowest cover - DHW timeseries
correlation levels, while reefs in 'high' clusters have the highest correlation levels
(@sfig-access-esm1-5-bioregion-dhwcor). Conversely, there are no consistent patterns across
bioregions in reef area-relative cover and mean timeseries DHW levels, with clusters having
similar mean DHW levels in many bioregions (@sfig-access-esm1-5-bioregion-dhwmean).
Similarly, there are no substantial differences between reef clusters in mean thermal
tolerance levels @sfig-access-esm1-5-bioregion-dhwtol). There are no consistent differences
across bioregions and GCMs between clusters in their reefs' initial proportion of habitable
area occupied by corals (@sfig-access-esm1-5-bioregion-initialproportion). Across most
bioregions there are not consistent differences between clusters in connectivity metrics
such as source-to-sink connectivity ratio (@sfig-access-esm1-5-bioregion-logsosi) and total
connectivity strength (@sfig-access-esm1-5-bioregion-totalconnectivity).
